---
title: "Chinese Investments and Vaccines"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "Chinese Vaccines and Investment"
author: "Advait Arun & Alex Lin"
date: "11/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Research Question

### Do Chinese vaccine distribution patterns track recent Chinese investment patterns?
The Chinese government is sending Chinese vaccines, made primarily by the companies Sinovac and Sinopharm, around the world. The news portrays most recipient countries as developing nations, many of which aren't able to pay for Western-made Pfizer and Moderna vaccines on their own. The United States government sometimes portrays Chinese vaccine donations as self-serving "vaccine diplomacy" that will end up boosting China. If this is true, China would be sending vaccines to its economic and political allies to bolster goodwill and to reinforce their dependence on China. 

#### Hypothesis: We expect Chinese vaccine disbursement to be higher in countries with higher amounts of Chinese investment.
For this purpose, we've used Chinese investment over the past five years (since 2016).

## Data
For Chinese investment, we got our data from the American Enterprise Institute. It goes all the way back to 2000 and includes lots of industry classifications. It also labels whether or not investment recipient countries are part of China's Belt and Road Initiative.

For Chinese vaccines, we got our data from Bridge Consultants, a Beijing-based economic development consultant group. They maintain a Chinese vaccine tracker, and we owe them a debt of gratidute for allowing us to use their data. All vaccine data is as of October 26th.

Later, we look at US Investment Data since 2016. Our data, also segregated by industry, comes from the Bureau of Economic Analysis.

## Data Preparation

```{r prep1, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(maps)
library(mapproj)
library(lubridate) #access lubridate to modify dates and times
library(biscale)
library(sf) # spatial data handling
library(sp)
library(cowplot)
library(ggthemes)
library(gridExtra)
library(RColorBrewer)

full_data <- read_csv("investments_vaccines_fulldata.csv")
data <- filter(full_data, Country != "China")
data <- data[,-1]

data$Region <- as.factor(data$Region)
data$Industry <- as.factor(data$Industry)
data$Subcategory <- as.factor(data$Subcategory)

data$quantity[is.na(data$quantity)] <- 0
data$total_doses_delivered[is.na(data$total_doses_delivered)] <- 0
data$doses_per_million[is.na(data$doses_per_million)] <- 0
data$quantity_adj[is.na(data$quantity_adj)] <- 0
data$usa_quantity[is.na(data$usa_quantity)] <- 0
data$usa_quantity_adj[is.na(data$usa_quantity_adj)] <- 0

#CREATING BI_CLASSES
bidata <- custom_bi_class(data, x = quantity_adj, y = doses_per_million, style = "quantile", dim = 5, include_lowest = T)
bidata <- rename(bidata, Investment_bi_Vaccine = bi_class)

bidata <- custom_bi_class(bidata, x = usa_quantity_adj, y = doses_per_million, style = "quantile", dim = 5)
bidata <- rename(bidata, USInvestment_bi_Vaccine = bi_class)

bidata <- custom_bi_class(bidata, x = quantity_adj, y = usa_quantity_adj, style = "quantile", dim = 5)
bidata <- rename(bidata, Investment_bi_USInvestment = bi_class)

bidata <- custom_bi_class(data, x = quantity, y = total_doses_delivered, style = "quantile", dim = 5, include_lowest = T)
bidata <- rename(bidata, Investment_bi_Vaccine_unadj = bi_class)

#MAP PREP
world_map <- map_data("world")
biworld <- full_join(bidata, world_map, by=c("Country"="region"))
```

Our FullData spreadsheet combines Chinese investment data, Chinese vaccine disbursement data, and US investment data. We use bi_class to create bivariate class mappings for each pair of the three variables in order to create bivariate chloropleths. We also have each variable in per-million-people format: Chinese vaccines per million people in recipient country, Chinese investment per million people, and US investment per million people. Finally, we join that spreadsheet with the world map.

There's one issue with our bivariate classing. Normally, the bi_class function breaks up observations into quantiles. But because our data has so many zeros, we had to break up our observations differently, and "quantile" did not work for the style parameter. We were able to circumvent this problem by creating our own bi_class functions, which can be found in the custom_bi_functions.R file. 

Our pre-cartogram data looks like this.

```{r data}
head(bidata)
```

## Bivariate Chloropleths

### Do Chinese Vaccines Follow Chinese Investment?

```{r biworldmap_unadj}
biworldmap1 <- ggplot() + geom_polygon(data=biworld, aes(x=long, y=lat, group=group, fill=Investment_bi_Vaccine_unadj), color="grey50", show.legend = FALSE, size = 0.1) +
  coord_map("albers", parameters=c(0,0)) +
  bi_scale_fill_custom(pal = custom_palette, dim = 4) + bi_theme() +
  labs(x="", y="", title="Chinese Investment since 2016") +
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), axis.ticks.x = element_blank(),axis.text.x = element_blank()) +
  labs(title = "Do Chinese Vaccines follow Chinese Investment?")
legend1 <- bi_legend_custom(pal = custom_palette,
                    dim = 4,
                    xlab = "Higher Investment ",
                    ylab = "Higher Vaccine Disbursement",
                    size = 8)
finalPlot1 <- ggdraw() +
  draw_plot(biworldmap1, 0, 0, 1, 1) +
  draw_plot(legend1, 0, .0, 0.4, 0.4, scale = 1)
finalPlot1
```

One problem with this visualization is that it potentially emphasizes Chinese vaccine donations and investments in larger countries over those in smaller countries. To adjust the data, we joined our existing dataset with population and GDP statistics for each country in 2020, taken from the World Bank. The new bivariate choropleth is as follows: 

```{r biworldmap}
biworldmap2 <- ggplot() + geom_polygon(data=biworld, aes(x=long, y=lat, group=group, fill=Investment_bi_Vaccine), color="grey50", show.legend = FALSE, size = 0.1) +
  coord_map("albers", parameters=c(0,0)) +
  bi_scale_fill_custom(pal = custom_palette, dim = 4) + bi_theme() +
  labs(x="", y="", title="Chinese Investment since 2016") +
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), axis.ticks.x = element_blank(),axis.text.x = element_blank()) +
  labs(title = "Do Chinese Vaccines follow Chinese Investment?")
legend2 <- bi_legend_custom(pal = custom_palette,
                    dim = 4,
                    xlab = "Higher Investment ",
                    ylab = "Higher Vaccine Disbursement",
                    size = 8)
finalPlot2 <- ggdraw() +
  draw_plot(biworldmap2, 0, 0, 1, 1) +
  draw_plot(legend2, 0, .0, 0.4, 0.4, scale = 1)
finalPlot2
```

Several pieces of insight can be extracted from this visualization. First, when adjusted for GDP and population, China invests aggressively and doles out large doses of vaccines in Southeast Asia and Latin America. Moreover, contrary to popular perceptions, Chinese presence in Africa is an inconsistent patchwork of investments and vaccine donations, but few countries receive high levels of both. Finally, Chinese presence in the West is almost entirely dominated by investments. To more closely examine regional trends and nuances, we have reconstructed the bivariate choropleths after isolating South and Southeast Asia, as well as Latin America. 

### Do Vaccines Follow Investment in South and Southeast Asia?

```{r bisseasiamap}
#creating new bi_classes after filtering the data
sseasia <- filter(data, Region == "SE Asia" | Region == "South Asia")
sseasia <- left_join(sseasia, world_map, by = c("Country" = "region"))
bi_sseasia <- custom_bi_class(sseasia, x = quantity_adj, y = doses_per_million, style = "quantile", dim = 4, include_lowest = T)
bi_sseasia <- rename(bi_sseasia, Investment_bi_Vaccine = bi_class)

#creating the visualization
biregionalmap <- ggplot() + geom_polygon(data=bi_sseasia, aes(x=long, y=lat, group=group, fill=Investment_bi_Vaccine), color="grey50", show.legend = FALSE, size = 0.1) +
  coord_map("albers", parameters=c(0,0)) +
  bi_scale_fill_custom(pal = custom_palette, dim = 4) + bi_theme() +
  labs(x="", y="", title="Chinese Investment since 2016") +
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), axis.ticks.x = element_blank(),axis.text.x = element_blank()) +
  labs(title = "Do Chinese Vaccines follow Chinese Investment?")
legend3 <- bi_legend_custom(pal = custom_palette,
                     dim = 4,
                     xlab = "Higher Investment ",
                     ylab = "Higher Vaccine Disbursement",
                     size = 8)
finalPlot3 <- ggdraw() +
  draw_plot(biregionalmap, 0, 0, 1, 1) +
  draw_plot(legend3, 0, .0, 0.4, 0.4, scale = 1)
finalPlot3
```

There's a lot of colors at play here, and it seems as if countries such as India and Afghanistan don't get a lot of either vaccines or investments from China when adjusted for GDP and population. Vietnam gets investments but little vaccines, while Thailand and Papua New Guinea get vaccines but relatively fewer investors. Meanwhile, the rest of the countries in the region, which are more or less friendly with China, get a good amount of both. Ultimately, because of the incredibly diverse vaccine distribution and investments landscape, a bivariate choropleth can become rather difficult to decipher. 

Let's take a look at Latin America. 

### Do Vaccines Follow Investment in Latin America and the Caribbean?

```{r bilacmap}
#creating new bi_classes after filtering the data
lac <- filter(data, Region == "Latin America")
lac <- left_join(lac, world_map, by = c("Country" = "region"))
bi_lac <- custom_bi_class(lac, x = quantity_adj, y = doses_per_million, style = "quantile", dim = 4, include_lowest = T)
bi_lac <- rename(bi_lac, Investment_bi_Vaccine = bi_class)

#creating the visualization
biregionalmap <- ggplot() + geom_polygon(data=bi_lac, aes(x=long, y=lat, group=group, fill=Investment_bi_Vaccine), color="grey50", show.legend = FALSE, size = 0.1) +
  coord_map("albers", parameters=c(0,0)) +
  bi_scale_fill_custom(pal = custom_palette, dim = 4) + bi_theme() +
  labs(x="", y="", title="Chinese Investment since 2016") +
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), axis.ticks.x = element_blank(),axis.text.x = element_blank()) +
  labs(title = "Do Chinese Vaccines follow Chinese Investment?")
legend3 <- bi_legend_custom(pal = custom_palette,
                     dim = 4,
                     xlab = "Higher Investment ",
                     ylab = "Higher Vaccine Disbursement",
                     size = 8)
finalPlot3 <- ggdraw() +
  draw_plot(biregionalmap, 0, 0, 1, 1) +
  draw_plot(legend3, 0, .0, 0.4, 0.4, scale = 1)
finalPlot3
```

Chile seems like a big winner here when we adjust for population and GDP. However, some other visualizations might make regional trends even more salient. 

### Bivariate Maps Don't Do the Trick
Ultimately, we've made a mistake using bivariate maps. Let's just try univariate maps in these regions to see if we can figure anything out. 

```{r unisseasiamap}
sseasia_investment <- ggplot() + 
  geom_polygon(data=sseasia, aes(x=long, y=lat, group=group, fill=quantity_adj), color="grey50") +
  coord_map("albers", parameters=c(0,0)) +
  scale_fill_gradient(name="Investment",low="whitesmoke",high="gold") +
  labs(x="", y="", title="Chinese Investment since 2016") +
  theme_void() +
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), axis.ticks.x = element_blank(),axis.text.x = element_blank())

sseasia_vaccines <- ggplot() + 
  geom_polygon(data=sseasia, aes(x=long, y=lat, group=group, fill=doses_per_million), color="grey50") +
  coord_map("albers", parameters=c(0,0)) +
  scale_fill_gradient(name="Vaccines",low="whitesmoke",high="darkred") +
  labs(x="", y="", title="Chinese Vaccine Distribution") +
  theme_void() +
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), axis.ticks.x = element_blank(),axis.text.x = element_blank())

lac_investment <- ggplot() + 
  geom_polygon(data=lac, aes(x=long, y=lat, group=group, fill=quantity_adj), color="grey50") +
  coord_map("albers", parameters=c(0,0)) +
  scale_fill_gradient(name="Investment",low="whitesmoke",high="gold") +
  labs(x="", y="", title="Chinese Investment since 2016") +
  theme_void() +
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), axis.ticks.x = element_blank(),axis.text.x = element_blank())

lac_vaccines <- ggplot() + 
  geom_polygon(data=lac, aes(x=long, y=lat, group=group, fill=doses_per_million), color="grey50") +
  coord_map("albers", parameters=c(0,0)) +
  scale_fill_gradient(name="Vaccines",low="whitesmoke",high="darkred") +
  labs(x="", y="", title="Chinese Vaccine Distribution") +
  theme_void() +
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), axis.ticks.x = element_blank(),axis.text.x = element_blank())

grid.arrange(sseasia_investment, sseasia_vaccines)
grid.arrange(lac_investment, lac_vaccines)
```

These two graph sets are illuminating. Comparing two univariate graphs has been easier for us than exploring one bivariate graph.

### South and Southeast Asia
Pakistan correctly gets a decent amount of Chinese investment and a lot of Chinese vaccines. Indonesia still gets the most of both. True to reality, India gets no Chinese vaccines but gets more Chinese investment. Interestingly enough, Lao PDR gets a lot of investment but has no reported vaccine disbursements. Meanwhile the rest of mainland Southeast Asia seems to have more vaccines than 

### Latin America
Mexico, the U.S.'s neighbor, gets Chinese vaccines and some investment. It's the same for Argentina. Brazil and Chile gets lots of both. Chile, one of the most purple countries in the bivariate map, gets fewer vaccines than Argentina does, but more investment than Argentina does.

## Breaking down the sectors of Chinese investment by region 

```{r investmentsectors}

library(RColorBrewer)

investments <- read_csv("Chinese Investments.csv")

#breaking down the investment sectors

cut_off_year = 2016
investments_new <- filter(investments, Year >= cut_off_year)
investments_new$`Quantity in Millions` <- gsub("\\$", "", investments_new$`Quantity in Millions`)
investments_new$`Quantity in Millions` <- gsub("\\,", "", investments_new$`Quantity in Millions`)
investments_new$`Quantity in Millions` <- as.numeric(investments_new$`Quantity in Millions`)

investments_new <- mutate(investments_new, Country=fct_recode(Country,
                                                              "UK" = "Britain",
                                                              "Palestine" = "West Bank and Gaza",
                                                              "Russia" = "Russian Federation",
                                                              "Antigua" = "Antigua and Barbuda",
                                                              "Bosnia and Herzegovina" = "Bosnia",
                                                              "Republic of Congo" = "Congo",
                                                              "Trinidad" = "Trinidad and Tobago",
                                                              "Syria" = "Syrian Arab Republic",
                                                              "Kyrgyzstan" = "Kyrgyz Republic",
                                                              "United Arab Emirates" = "UAE",
                                                              "Ivory Coast" = "Côte d'Ivoire"))
                                                                                            
investments_new <- select(investments_new, -Region)

investments_by_region <- left_join(investments_new, data_adj, by = "Country")

investments_by_region$Region[investments_by_region$Region=="Arab Middle East and North Africa"] <- "MENA"

nb.cols <- 15
mycolors <- colorRampPalette(brewer.pal(8, "Accent"))(nb.cols)

ggplot(data = investments_by_region) + geom_bar(aes(x = Region, y = `Quantity in Millions`, fill = Sector), stat = "identity", position = "fill") + 
  scale_fill_manual(values = mycolors)

```


## Does BRI status boost either vaccine distribution or investments from China? 

```{r BRI}

```

## So, Do Vaccines Follow the Money?
